FraCaS: Temporal Analysis

### 2020-06-09 paper2020/paper.tex
example

\[\begin{array}{l}
∀x. novel(x) ∧ \\
∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,x,t_3,t_4) \\
⟶ ⊥ \\
\end{array}
\]
Note here that the scope for the existential is extended beyond the
scope of \textbf{P1}, and its polarity switched (to universal). This extension can work according to
\citet{unger_dynamic_2011}, and our implemented analysis of
anaphora\ifanon (citation anonymized)
\else\citep{bernardy_computational_2018,bernardy_wide-coverage_2019}\fi.

@@

{\small
\begin{verbatim}
s_279_1_p=
 sentence (useCl past pPos
  (predVP (usePN (lexemePN "smith_PN"))
          (advVP (complSlash (slashV2a (lexemeV2 "write_V2")))
                     (detCN (detQuant indefArt numSg) (useN (lexemeN "novel_N"))))
                     ((lexemeAdv "in_1991_Adv")))))
s_279_3_h=sentence (useCl past pPos (predVP (usePN (lexemePN "smith_PN")) (advVP (complSlash (slashV2a (lexemeV2 "write_V2")) (usePron it_Pron)) ((lexemeAdv "in_1992_Adv")))))
\end{verbatim}
Interpretation by our dynamic and temporal semantics:

\begin{verbatim}
FORALL (fun a=>novel_N a)
(fun a=>(exists (b: Time),
((exists (c: Time),
(IS_INTERVAL Date_19910101 b /\
 IS_INTERVAL c Date_19911231 /\
 IS_INTERVAL b c /\
 appTime b c (write_V2 a) (PN2object smith_PN))))) ->
Not (exists (f: Time),
((exists (g: Time),
 (IS_INTERVAL Date_19920101 f /\
  IS_INTERVAL g Date_19921231 /\
  IS_INTERVAL f g /\
  appTime f g (write_V2 a) (PN2object smith_PN)))))).
\end{verbatim}
Unfolding definitions by Coq:

\begin{verbatim}
forall x : object,
novel_N x ->
(exists b c : Z,
   Date_19910101 <= b /\
   c <= Date_19911231 /\
   b <= c /\ write_V2 x SMITH b c) ->
(exists f g : Z,
   Date_19920101 <= f /\
   g <= Date_19921231 /\
   f <= g /\ write_V2 x SMITH f g) ->
False
\end{verbatim}
Proof:

\begin{verbatim}
Theorem  problem279 : Problem279aFalse.
cbv.
intros novel isSmithsNovel P1 H.
destruct P1 as [t0 [t1 [ct1 [ct2 [ct3 P1]]]]].
destruct H as [u0 [u1 [cu1 [cu2 [cu3 H]]]]].
specialize writeUnique with (x := novel)(y := SMITH) as A.
unfold UniqueActivity in A.
specialize (A _ _ _ _ P1 H) as B.
lia.
Qed.
\end{verbatim}
}

>>>>

\begin{equation}
  \label{eq:smith-novel-1}
\begin{array}{l}
∀x. novel(x) ∧ \\
∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,x,t_3,t_4) \\
⟶ ⊥ \\
\end{array}
\end{equation}
Note here that the scope for the existential is extended beyond the
scope of \textbf{P1}, and its polarity switched (to universal). This extension can work according to
\citet{unger_dynamic_2011}, and our implemented analysis of
anaphora\ifanon (citation anonymized)
\else\citep{bernardy_computational_2018,bernardy_wide-coverage_2019}\fi.

@@

{\small
\begin{verbatim}
s_279_1_p=
sentence 
(useCl past pPos
 (predVP
   (usePN (lexemePN "smith_PN"))
   (advVP
     (complSlash
      (slashV2a (lexemeV2 "write_V2")))
     (detCN (detQuant indefArt numSg)
       (useN (lexemeN "novel_N"))))
       (lexemeAdv "in_1991_Adv")))
s_279_3_h=
sentence
(useCl past pPos
 (predVP (usePN
   (lexemePN "smith_PN"))
   (advVP
     (complSlash
      (slashV2a (lexemeV2 "write_V2"))
      (usePron it_Pron))
     (lexemeAdv "in_1992_Adv"))))
\end{verbatim}
Interpretation by our dynamic and temporal semantics:

\begin{verbatim}
FORALL (fun a=>novel_N a)
(fun a=>(exists (b: Time),
((exists (c: Time),
(IS_INTERVAL Date_19910101 b /\
 IS_INTERVAL c Date_19911231 /\
 IS_INTERVAL b c /\
 appTime b c (write_V2 a)
   (PN2object smith_PN))))) ->
Not (exists (f: Time),
((exists (g: Time),
 (IS_INTERVAL Date_19920101 f /\
  IS_INTERVAL g Date_19921231 /\
  IS_INTERVAL f g /\
  appTime f g (write_V2 a) 
    (PN2object smith_PN)))))).
\end{verbatim}

Unfolding definitions at the Coq level:
\begin{verbatim}
forall x : object,
novel_N x ->
(exists b c : Z,
   Date_19910101 <= b /\
   c <= Date_19911231 /\
   b <= c /\ write_V2 x SMITH b c) ->
(exists f g : Z,
   Date_19920101 <= f /\
   g <= Date_19921231 /\
   f <= g /\ write_V2 x SMITH f g) ->
False
\end{verbatim}
(Very close to \cref{eq:smith-novel-1})

Proof:
\begin{verbatim}
Theorem  problem279 : Problem279aFalse.
cbv.
intros novel isSmithsNovel P1 H.
destruct P1 as
   [t0 [t1 [ct1 [ct2 [ct3 P1]]]]].
destruct H as
   [u0 [u1 [cu1 [cu2 [cu3 H]]]]].
specialize writeUnique
  with (x := novel)(y := SMITH) as A.
unfold UniqueActivity in A.
specialize (A _ _ _ _ P1 H) as B.
lia.
Qed.
\end{verbatim}
}

Critical: UniqueActivity; lia does arithmetic work.
